Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-16 paper-2023-acsos-pulverization.tex
Intro: adjusting the incipits, minors, and +toc

\newcommand{\ourframework}{\texttt{pulse-refrakt}}

@@

Recent research trends in topics related to the Internet of Things promote the idea of a cloud-edge continuum~\cite{DBLP:journals/iot/BittencourtISFM18},
an infrastructure composed of heterogeneous devices ranging from large and powerful cloud servers to small connected things
on which software computes in a distributed fashion,
optimising for performance and resource utilisation.
%
This kind of infrastructure is ideal for the execution of \ac{cas},
which have components that,
depending con the conditions at hand,
could benefit from being deployed on different devices.
%
Exploiting this intriguing infrastructure, however, poses new challenges to the application designer.
%
Commonly, in fact, applications are designed with a specific infrastructure in mind,
whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
unless captured and encapsulated away from it:
the higher the coupling between the application and the infrastructure,
the harder it is for the former to exploit the latter's full potential and adapt to changes.

Previous work suggests a technique for decoupling business logic and deployment concerns
called \emph{pulverization}~\cite{FI2020-pulverization}.
%
In short, the core idea is to consider devices as \emph{logical} entities,
whose software can be designed as (or broken down to, if already existing) an ensemble of five micro-components
(behaviour, state, communication, sensors, and actuators)
whose actual host can be decided at deployment time.
%
This way, the designer can focus on the business logic at hand,
delaying an optimised deployment to later stages of the sofware design,
thus gaining resilience to changes in the infrastructure or in non-functional requirements.
%
One limitation of the current pulverisation approach regards dynamicity:
although the application can be deployed on arbitrary systems,
there is no support for the reconfiguration of components in face of changing conditions or requirements.

In this work,
we continue the research on pulverised systems by providing two key contributions:
\textbf{first}, we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
\textbf{second}, we provide a practical implementation called \ourframework{} which,
to the best of our knowledge,
is the first \emph{practical framework supporting pulverisation} for a general-purpose language.
%
We exercise \ourframework{} with two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.

\begin{itemize}
    \item need for deployment-independent specs
    \item need for runtime reconfiguration (achieve \ac{qos} in face of changing conditions --- green computing?)
    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
\end{itemize}

@@

\section{Related Work}

>>>>

\newcommand{\ourframework}{\textsc{pulse-refrakt}}

@@

Recent technological and scientific advances
 are extending 
 both the \emph{kinds of applications and systems} that are being addressed  
 as well as their \emph{supporting infrastructure}.
%
Emerging applications
 are, for instance, those based on \emph{\acp{cas}}~\cite{DBLP:journals/sttt/NicolaJW20},
 namely collections of devices and agents
 that interact to solve problems or provide services while adapting as a whole to dynamic environments.
%
Examples include 
  large-scale sensor-actuator networks, 
  swarms of robots,
  social computing systems,
  crowds of wearable-augmented people, etc.
  supporting activities like monitoring,
  transportation, coordination, and collective intelligence.
%
Regarding infrastructure,
 it is mounting the idea of the \emph{cloud-edge continuum}~\cite{DBLP:journals/iot/BittencourtISFM18}: 
a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things})
on which software can compute, store, and exchange data in a distributed fashion, while 
optimising for performance and resource utilisation.
%

This kind of infrastructure is also valuable for  \acp{cas}~\cite{DBLP:journals/tpds/HongCHGZ19,DBLP:journals/comsur/WangZZMLW20,DBLP:journals/comsur/AfrinJRRWH21,IEEE-IoTJ-pulverization-simulation},
which generally feature components that,
depending on the conditions at hand,
could benefit from being deployed on different devices or from \emph{offloading} some of their  tasks.
%
Exploiting this %intriguing infrastructure, 
 however, poses new challenges to \ac{cas} application designers.
%
Commonly, in fact, applications are designed with a specific infrastructure in mind,
whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
unless captured and encapsulated away from it:
the higher the coupling between the application and the infrastructure,
the harder it is for the former to exploit the latter's full potential and adapt to changes.

Previous work suggests a technique for decoupling business logic and deployment concerns
in \acp{cas} called \emph{pulverisation}~\cite{FI2020-pulverization}.
%
In short, the core idea is to consider devices as \emph{logical} entities,
whose software can be designed as (or broken down to, if already existing) an ensemble of \emph{five components
(behaviour, state, communication, sensors, and actuators)}
 that can be deployed with flexibility on available infrastructure without (ideally) affecting application functionality. %whose actual host can be decided at deployment time.
%
This way, the designer can focus on the business logic at hand,
delaying an optimised deployment to later stages of the sofware design,
thus gaining resilience to changes in the infrastructure or in non-functional requirements.
%
One limitation of the current pulverisation approach regards dynamicity:
although the application can be deployed on arbitrary systems,
there is no support for the reconfiguration of components at runtime, as a reaction to changing conditions or requirements.

In this work,
we continue the research on pulverised systems by providing two key contributions:
\textbf{first}, we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
\textbf{second}, we provide a practical implementation in Kotlin called \ourframework{} which,
to the best of our knowledge,
is the first \emph{practical framework supporting pulverisation}. % for a general-purpose language.
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.

\meta{
\begin{itemize}
    \item need for deployment-independent specs
    \item need for runtime reconfiguration (achieve \ac{qos} in face of changing conditions --- green computing?)
    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
\end{itemize}
}

The manuscript is organised as follows.
%
\Cref{sec:background} provides background on \meta{deployment and} pulverisation.
%
\Cref{sec:contribution} describes the proposed \ac{dsl} and platform.
%
\Cref{sec:evaluation} provides an evaluation of the approach.
%
\Cref{sec:rw} covers related work.
%
Finally, \Cref{sec:conclusion} concludes the paper and highlights directions for future work.

@@

\section{Related Work}
\label{sec:rw}
